\articlehead{Furry Impressions}{Rabbit}{2013}

There are always key moments in any human relationship, whether said relationship is rooted in business, romance, politics, or pretty much anything else. When I was an adolescent, one of the most current memes in society was that a person's first impression of someone or something was the most crucial moment of all. While of course I can't recall all this in encyclopedic detail, at that time the market was flooded with books on how to improve your first impression, and said books were filled with charts ``proving'' just how vitally important this was to success in life's endeavors. The principle was even carried over into academics—kids were given ``fun math'' to do on their first day in school, to improve that vital initial impression. I recall this pretty well because, being a teen at the time, it provided me with my first-impression of the self-help book industry and, well\ldots We all know how lasting a first impression can be, no?

At any rate, it's inevitable that furs coming into the fandom tend to undergo a whole series of ``firsts''. One of these, which was often life-changingly profound to people of my own age, was the first discovery that there were other people as crazily in love with anthropomorphic art and stories and costumes and such as we were. Perhaps the power of this moment has lessened with time; in my own case I discovered the fandom at an age in excess of 35 because of the simple fact that until then there was no Internet access where I live. After over three decades of isolation, well\ldots I was practically turning somersaults with joy! I was also even more socially awkward than I am now and had no idea whatsoever how to handle myself online at all, much less among furs. A handful of individuals (to whom I remain eternally grateful) helped me along, were patient when I wrote them too-long and too-personal e-mails, etc. These people formed my first impression of the fandom; had they brushed me off it would've just about broken me.

My next ``first'' was meeting other furs in person. In my case I met two on the same night. It was their first time too, and clearly we were more than a little scared of each other. But the event was a success overall, mostly because we all worked at making it so. More happy memories, more growth as a fur.

After that, there was really only one ``first'' left, and that was my first furmeet. It happened to be Mephit—I'm not certain but I think it was \#3 (counting the infamous pizza party as \#1). That experience really pushed my limits, but it also made me certain that I wanted to be part of this fandom in the long term.

While I didn't sit down to write an autobiography—my original planned title was ``Going to your First Furcon''—the more I remembered my early days as a fur. Which in turn reminded me of how tremendously grateful I am to those who held my hand, who listened to me prattle, who put up with my poor social skills and generally made me fit to become part of this wonderful society. I've always tried to remember how much I owe these folks—some of whom I'd be embarrassed to meet today—and have attempted to ``pay it forward'' by taking the time to chat with newcomers to the irc channel I frequent and actively try to get to know lost-looking people wandering around the corridors at cons. Because they are me, you see—me as I was, and me as I'd have remained had I decided that I couldn't possibly fit in with the furry crowd and ought to go off and be alone again. Sure, sometimes it backfires—the chemistry can't always be right. But I try to help, and it's through my efforts that I honor those who helped me.

So, here's my challenge to both of the readers who I expect will still be with me at this point in the article\ldots

Remember that you too once wandered the chatrooms and convention halls as an awkward misfit, if only because you were a teenager and all teens are awkward misfits. Remember that you were too shy to hug the fursuiter, and sat and ate your lunch all by yourself because no one sat down at your table with you. Remember that its entirely possible to be lonely in a crowd; indeed, it's inevitable until someone lets you in and makes you part of the crowd. Remember that you sat in the panel afraid to raise your hand and ask what you really wanted to know. And as you remember these things and see them reflected in the eyes of a stranger standing by himself in a corridor and looking a little dazed, remember that he's forming his first impressions of our fandom. His most powerful impressions, in fact. The ones that he'll carry with him and judge us by forevermore.

I was damned lucky to get such a wonderful first impression of furdom. Or perhaps not—as I said, numerous individuals went far out of their way on my behalf. But at every furmeet and convention you ever go to, there are first impressions being formed all around you at every moment of every hour. I do my best to make them good ones, as I've said, as much as anything in an effort to honor those who did the same for me.

Do you owe the fandom any less than I do?
