\articlehead{Not-So-Distant Cousins}{Rabbit}{2013}

\textit{This is a lightly-edited reprint of a column in Anthro Magazine that first appeared in Issue \#14, in 2007.}

Perhaps my favorite activity at conventions is having dinner with groups of friends at local eateries. Anyone who knows me well will recognize that I'm pretty fond of my chow to begin with, and to be able to share my dining experience with a (usually) mixed group of old friends and new acquaintances is, well, the highlight of my calendar. Usually, at least once during a con I'll try and round up a suitable group, and off we go for what is always a memorable time out.

One of the most remarkable such con-dinners I've had in recent years took place in Memphis, during Mephit, at the Germantown Commissary. A group of about fifteen of us of mixed ages and of varying degrees of my acquaintanceship formed a convoy and ran across town to this trendy establishment, which allegedly sold the best barbeque in town. On the way our convoy broke up and some of us were separated, causing much anxiety. The Commissary proved to be the most highly-overrated restaurant I've ever experienced; it was crowded, the servers were rude to the point of surliness, and the food was okay at best. At any other time I'd have been very unhappy with my evening.

But because of the people I was with, the experience was absolutely magic.

I don't know what it is about fen (a common slang-term for SF and other geeky-type fandom members), but within minutes of sitting down we were off and running as if we did this sort of thing every evening of our lives. I can't claim to recall everything we discussed; the conversation was much too witty, free-ranging and articulate for me to remember it all. I do know that we went from ribald jokes to the nature of the universe and back several times in the blink of an eye. It was, in other words, the kind of stimulating and intelligent conversation that I spend most of my life starved for—that most of us fen, I suspect, spend our everyday lives starved for. Certainly, it was for me rain to a desert.

At one point, while we were discussing the finer points of terraforming, a good friend of long standing who's also very interested in high-tech and has a degree from a prestigious university interrupted. ``Wait a minute!'' he commented. ``We're furs! We're not supposed to be talking about this kind of stuff! Aren't we supposed to be debating which part of the fox is the floofiest?''

The comment got a good laugh, but it also made me think. My friend was right, in that the dinner conversation at this particular meal had a very definite SF flavor. Even more, I realized, the same was true of most of the similar dinners I've enjoyed for all these many years now. We're far more likely to discuss artificial intelligence at these affairs than vulpine floofiness, for example. And while fur-related subjects like gengineering probably get more than their share of air-time, even these conversations wouldn't be out of place at, say, WorldCon. In other words, while the furry fandom did in fact (as near as I can tell, not having been around at the time) split off from the world of SF, there's still an awful lot of SF left in us even these many years down the road. While this may not be true of all or even most fur-fans, I do have definite preferences in mind when selecting my dinner-companions, after all, and the ability to discuss a wide variety of subjects both amicably and intelligently ranks high among them—there's certainly enough of us SF-derived types around to give the fandom a definite and recognizable flavor.

I had this brought home even more forcefully to me at Rain Furrest one year, where I (being one of the guests of honor) interacted quite a bit with the con chairman. He was strictly an SF fan, not a fur, and the local furs had drafted him to run their con because at that time no one else thought they had enough experience. He and I had a quite pleasant conversation about the differences between a furmeet and an SF con… and, frankly, there weren't many. Furs consume less alcohol, as a rule, and require a little more in the way of costuming support. But all the major elements are pretty much the same, derived almost entirely from the SF cultural parent. Even much of the lingo is the same; I didn't blink when he used the term `SMOF' (Secret Master of the Fandom), and he recognized my use of `fen' and `mundane' (meaning anyone other than a fen) right off. I don't recall if the terms actually came up, but if he'd told me that a mutual acquaintance had `gafiated' (Gotten Away From It All) or `fafiated' (been Forced Away From It All), I'd have understood him instantly. In other words, furs and SF fen share a private language.

I suppose I can understand how the split between furs and mainstream SF fans came to be. There are distinct cultural differences, and probably most furs aren't as hard-science driven as the SF fandom. Even so, there's enough of us tech-rooted types around that I've never had problems putting together dinner-groups at fur-meets which I'd match, man-for-man, against just about any similar SF-based group in terms of general geekiness and science-savvy. It's a shame we've grown apart, in some ways, and even more a shame that in some circles there seems to be actual animosity between the two groups. We may be two distinct cultures, but we're at least as alike as, say, Canadians and US citizens. We're non-mundanes, in a world where non-mundanes are all too rare and precious. Therefore we should be treating each other as respected and beloved cousins, if not brothers and sisters. Not calling each other names, as I regret to report happens all too often.

We're all fen together, is what I'm trying to say. Natural allies, not rivals. I mean, how many places can you find people who not only enjoy discussing terraforming over barbeque, but are good at it? Not many, in this sad and intellectually-declining world. And, while I can't speak for anyone but myself, I fully intend to treasure such individuals wherever I may find them. Whether that be in a fursuit, wearing an Imperial Storm Trooper outfit, encased in a suit of armor, or behind a deck of Magic cards. We're fen, all of us. And there just aren't enough of us around to allow for snootiness when choosing friends.
