\articlehead{Full Cotton-Polyester Blend Jacket}{Katzenjammer}{2012}

\begin{quotation}
  When I was pulling together the contributions page, I did my best to think through all the types of submissions that would fit in here on [adjective][species].  I wound up tossing ``fiction'' in there just in case someone could think of a way to write meta-furry fiction.  I could think of one example by WhyteYote that would’ve fit, except for the fact that it was erotica.

  However, lo and behold, our first piece of fiction arrived just a few days ago. Katzenjammer, after hearing stories about overzealous convention security, pulled together a short piece about a rent-a-cop taking his job protecting the animal people a little too seriously\ldots

  --Ed.
\end{quotation}

``Why did you do this, Lewis?''

``The rules state only one handler per. The lady or gentleman involved was attempting to bring two. That is double the allotted number.''

``But did you have to tackle one of them?''

``They were not responding to my verbal direction.''

``People here are calling us Nazis. I don’t need to tell you that is not the image we are trying to project.''

``With all due respect ma’am but I’m certain that at the Nazi ‘furry’ convention we would not be having this conversation as no one would consider bringing two handlers into the headless lounge because they had a respect for order that is sadly lacking here.''

``Oh. \ldots are you some kind of World War II history buff, then?''

``Huh? \ldots no.''

``Oh. Well I’m just concerned because you’ve been doing such a great job so far. I mean, we all appreciate you taking this seriously. I know that this can be a job that is difficult to take seriously. But there are limits. How did you come across this job anyways?''

``Well I really don’t know much about your organization here, but I saw the posting at the temp agency for additional help and it seemed\ldots  it seemed important to me.''

``Important?''

``Yes.''

``That’s a very strange choice of words.''

``I apologize.''

Sigh.  ``So, how am I going to explain this to the staff?''

\secdiv

It was 1983 when we landed in the swamp.

They told me that the mouse lived here. That’s all I needed to know. Why did he live here, I wondered, in this oppressively sweltering land? Perhaps to keep his enemies at bay. Perhaps as a challenge to the righteous. I was righteous. Did he think that a lay over at the stagnant, labyrinthine Atlanta airport would stop me? Did he think that making the best option for lunch the airport Applebees would slow me down? There was a lot of suffering in that Applebees.

‘Muggy.’ That’s how people from the midwest describe a hellhole. People with values. Good people. That’s how my parents described it. They didn’t deserve this.

We arrived.

The sign had printed on it ‘The happiest place on earth’ but I understand that what it was saying was ‘Abandon all hope ye who enter here.’ The snow cones around every corner seemed like a sick joke. What chance did they have?

It wasn’t difficult to find him. The mouse. We stood in line, I would have my meeting, it was imminent. But wait.

They came out of nowhere but the incessant, shrill cackling was unmistakable. Teenagers! They descended on the mouse. I couldn’t tell what was happening and they disappeared as quickly as they had arrived.

When I finally got a view of what was going on I saw it happen. He just collapsed. The mouse. How could he be so weak. His image adorned every surface that could take it for fifty miles around. This was the reason I had come so far? Yes. This was nothing, a surprise attack by cowards, no real damage was done. A prank. My concerns were unwarranted. I thought.

But he stayed down. His vassals arrived.

Then his head came off.

What?

His HEAD came OFF.

WHAT.

They lifted it off, like it was the lid on a jar of pickles that a friend had already failed to open but regardless claimed credit for ‘totally loosening it’ when you succeeded.

I don’t. I don’t understand.

Yes. You do.

Oh god.

It’s not over yet.

Why.

It’s just not.

What is that inside of him? Is that a person? What was going on here? This was madness! Was I wrong? Was I not prepared? My mind was too fragile. Why didn’t they warn me?

But it wasn’t over, he was right. In this heat, this\ldots  interloper started\ldots  vomiting.

Alright that’s just gross.

I know. I’m sorry. But it happened. It seems so obvious. He was vomiting out the magic. It kept coming out of him until it stopped. Then, mercifully, they took him away.

But it stayed. It just lay there, baking in the sun. The magic. To say that it was ruined was offensive. It was destroyed. Annihilated. Extinguished.

Never again.

\secdiv

``Hey? Hey are you listening to me? You kind of zoned out there for a while. Just don’t it again, alright. Try to relax huh, try to have some fun. This isn’t exactly Fort Knox. You’re too tense.''
